No proof nets for \capsabbrev{mll} with units\\[5pt]\Large Proof equivalence in \capsabbrev{mll} is \capsabbrev{pspace}-complete

### 2014-04-03 main/LICS.tex
balanced header by removing address
added reviews, started making changes

\author{
  \IEEEauthorblockN{Willem Heijltjes}
  \IEEEauthorblockA{
  		University of Bath
	\\	Claverton Down
	\\	Bath BA2~7AY
	\\	w.b.heijltjes@bath.ac.uk
	}
\and
  \IEEEauthorblockN{Robin Houston}
  \IEEEauthorblockA{
		Kiln Enterprises
	\\	robin@kiln.it
	}
}

@@

Later, a striking technical innovation of linear logic was the introduction of proof nets \cite{Girard-1987,Danos-Regnier-1989}, which define a canonical form for proofs in the unitless fragment of multiplicative linear logic -- two proofs are equivalent if and only if they have the same proof net -- so proof nets offer a simple decision procedure for proofs in this fragment. This naturally raises the question whether proof nets can be extended to work in the presence of units. The work in this direction begins with \cite{Trimble-1994,Blute-Cockett-Seely-Trimble-1996} via \cite{Strassburger-Lamarche-2004} and perhaps culminates in \cite{HughesFreeStar,HughesMLLProofNets}; but these proof nets are not canonical and must be identified up to a rewiring equivalence.

@@

In this paper we establish that the proof equivalence problem for multiplicative linear logic with units is \capsabbrev{pspace}-complete. It follows that there can be no satisfactory canonical notion of proof net.

@@

\begin{proposition}[\cite{Danos-Regnier-1989}]
\label{prop:correctness and sequentialisation}
%
%
If $\Pi\toNet\links$ and $\Pi$ has conclusion $\Gamma$, then $\links$ is a proof net for $\Gamma$.
%For a proof $\Pi$ with conclusion $\Gamma$, if $\Pi\toNet\links$ then $\links$ is a proof net for $\Gamma$.
%
For a net $\links$ for $\Gamma$, there is a proof $\Pi$ of $\Gamma$ such that $\Pi\toNet\links$ (\emph{sequentialisation}).
%
\end{proposition}

@@

We will denote by $A^n$ the sequent consisting of $n$ occurrences of a formula $A$.
%
Given a sequent $\Gamma=A_1,\dotsc,A_n$ we will write $\bigtn\Gamma$ for $A_1\tn\cdots\tn A_n$, and $\bigparr\Gamma$ for $A_1\parr\cdots\parr A_n$.
%
In diagrammatic notation, a big disc will represent an $n$-ary tensor over only $\bot$-formulae, $\bigtn(\bot^n)$, and a big circle will represent an $n$-ary par (or a sequent) over only $1$-formulae, $\bigparr(\1^n)$.
%
A linking between $\bigtn(\bot^n)$ and $\bigparr(\1^n)$ will be represented by a wide arrow, as illustrated below.
\[
\begin{tikzpicture}[x=-5mm,y=5mm,octo]
	\node[big_] (A) at (-1,2) {};
	\node[big1] (!) at (-1,0) {};
	\node at (.5,1) {$\Rightarrow$};
	\draw (2,2) node[bullet] (a) {} -- (3,2) node[bullet] (b) {} -- (4,2) (5,2) -- (6,2) node[bullet] (c) {}; 
	\node[circ] (1) at (2,0) {}; \node[circ] (2) at (3,0) {}; \node[circ] (3) at (6,0) {};
	\draw[dotted] (4,2)--(5,2) (4,0)--(5.1,0);
	\begin{scope}[thepink,->]
			\draw[big>] (A) to (!.center);
			\draw[bend right=10] (a) to (1.center);
			\draw[bend right=10] (b) to (2.center);
			\draw[bend right=10] (c) to (3.center);
	\end{scope}
\end{tikzpicture}
\]
Two formulae $\bigtn(\bot^{i+1})$ and $\bigtn(\bot^{j+1})$ may together connect to a formula $\bigparr(\1^{i+j+1})$, as illustrated below for $i=2$ and $j=3$.
%
Nodes will be labelled $i'$ for $i+1$, so that both the abbreviated formulae and the arithmetic of connecting them remain intuitive.
%
\[
\vc{
  \begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\draw (0,0) node[bullet] (a) {}--
		  (1,0) node[bullet] (b) {}--
		  (2,0) node[bullet] (c) {};
	\draw (4,0) node[bullet] (d) {}--
		  (5,0) node[bullet] (e) {}--
		  (6,0) node[bullet] (f) {}--
		  (7,0) node[bullet] (g) {};
	\foreach \i in {1,...,6} {\node[circ] (\i) at (\i,3) {};}
	\begin{scope}[thepink,->]
			\draw[bend right=10] (a) to (1);
			\draw[bend right=10] (b) to (2);
			\draw[bend right=10] (c) to (3);
			\draw[bend left=10]  (d) to (3);
			\draw[bend left=10]  (e) to (4);
			\draw[bend left=10]  (f) to (5);
			\draw[bend left=10]  (g) to (6);
	\end{scope}
  \end{tikzpicture}
}
\quad\Rightarrow\quad
\vc{
  \begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\node[big_] (A) at (-1,0) {$2'$};
	\node[big_] (B) at ( 1,0) {$3'$};
	\node[big1] (!) at ( 0,3) {$5'$};
	\begin{scope}[thepink,->,big>]
		\draw[bend right=15] (A) to (!);
		\draw[bend left=15]  (B) to (!);
	\end{scope}
  \end{tikzpicture}
}
\]
%
A dashed edge will indicate that two vertices are connected, in a proof net or a switching graph, as illustrated below.
\[
\begin{tikzpicture}[x=5mm,y=5mm,octo]
	\node[bullet] (a) at (0,0) {}; \node[bullet] (b) at (2,0) {}; \draw[dashed] (a)--(b);
\end{tikzpicture}
\qquad
\begin{tikzpicture}[x=5mm,y=5mm,octo]
	\node[circ] (a) at (0,0) {}; \node[circ] (b) at (2,0) {}; \draw[dashed] (a)--(b);
	%\coordinate (a') at ($(a) + (.6,0)$); \coordinate (b') at ($(b) + (-.6,0)$);
	%\draw[thepink,dashed] (a')--(b'); \draw[thepink,->] (a')--(a.center); \draw[thepink,->] (b')--(b.center);
\end{tikzpicture}
\]

@@

The scopes of a par correspond to the possible subproofs of its introduction rule in a sequentialisation of the proof net that it occurs in.
%
In the graph of a proof net, the scope of a par $\named v\parr$ may be \emph{contracted} to a single vertex $v$ by removing all vertices except $v$,  and re-attaching the edges of any removed vertices to $v$.
%
In the example below, a scope of a (ternary) par is indicated (left) and contracted (right).
\[
\begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\draw (-2.25,0) node[bullet] (x) {} -- (-1.25,0) node[bullet] (y) {}
		  (.25,0) node[bullet] (a) {} -- (1.25,0) node[bullet] (b) {}
		  ( 2.75,0) node[bullet] (c) {} -- (3.75,0) node[bullet] (d) {};
	\node[circ] (0) at (5.75,0) {};
	\node[circ] (1) at (-1,3) {};
	\node[circ] (2) at (1,3) {}; \node[circ] (3) at (2,3) {}; \node[circ] (4) at (3,3) {};
	\draw[rounded corners] (0,2) rectangle (4,4);
	\draw (4,3) -- (5.5,3) node[bullet] (e) {};
	\begin{scope}[thepink,->]
			\draw[bend right=10] (x) to (1);
			\draw[bend right=20] (y) to (2);
			\draw[bend right=10] (a) to (2);
			\draw[bend right=10] (b) to (3);
			\draw[bend left=10]  (c) to (3);
			\draw[bend left=10]  (d) to (4);
			\draw[bend right=10] (e) to (0);
	\end{scope}
	\draw[dashed,rounded corners] (-.25,-.5) rectangle (4.25,4.25);
\end{tikzpicture}
\qquad\qquad
\begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\draw (-2.25,0) node[bullet] (x) {} -- (-1.25,0) node[bullet] (y) {};
	\node[circ] (0) at (3,0) {};
	\node[circ] (1) at (-1,3) {};
	\node[draw,rounded corners,minimum size=16pt] (box) at (1,3) {};
	\draw (box)--(3,3) node[bullet] (e) {};
	\begin{scope}[thepink,->]
			\draw[bend right=10] (x) to (1);
			\draw[bend right=20] (y) to (box);
			\draw[bend right=10] (e) to (0);
	\end{scope}
	\path (0,4.25)--(1,4.25);
\end{tikzpicture}
\]
%
The contraction of scopes may replace the switching condition as a correctness criterion.
%
The following is a variant of the local retraction algorithm by Danos \cite{Danos-1990}.

\begin{proposition}
\label{prop:scoping correctness}
A linking $\links$ for a sequent $\Gamma$ is a proof net if and only if each $\named v\parr$ is a port of a sub-sequent $s(v)\leq\Gamma$ such that:

\begin{enumerate}
	\item
sub-sequents either disjoint or strictly nested: if $s(v)\cap s(w)\neq\varnothing$ then $s(w)<s(v)$ or $s(v)<s(w)$; and
	\item
for each $\named v\parr$ the graph $\sigma(v)=(s(v),\links|_{s(v)})$ becomes a tree when all immediate sub-sequents are contracted.
\end{enumerate}
%
\end{proposition}

\begin{IEEEproof}
For the `if' direction, it follows by induction on the nesting of sub-sequents that each graph $\sigma(v)$ satisfies the switching condition.
%
For the `only if' direction, given a sequentialisation of $(\Gamma,\links)$, a sub-sequent $s(v)\leq\Gamma$ for each $\named v\parr$ is found by taking the conclusion $\Delta, A\named v\parr B$ of its introduction rule, below.
\[
	\infer[\MLLlabel p]{\Delta,A\named v\parr B}{\Delta,A,B}
\]
\vskip-12pt
\end{IEEEproof}

In a proof net, the \emph{kingdom} and the \emph{empire} of a vertex $v$ are respectively the smallest and largest subnet that have $v$ as a port.
%
In working with the rewiring relation, the notion of empire can be particularly useful.
%
Firstly, we repeat the following proposition.

\begin{proposition}[{\cite[Proposition 2.b]{Bellin-vandeWiele-1995}}]
\label{prop:empire propagation}
The empire $\sigma$ of $v$ is determined by propagation from $v$: 1) through links; 2) up towards subformulae; 3) into a tensor if one of its subformulae is in $\sigma$; 4) into a par if all its subformulae are in $\sigma$; 5) except never into a par or tensor if $v$ is a subformula.
\end{proposition}

@@

\begin{lemma}
\label{lem:rewiring within empire}
For a proof net $(\Gamma,\links)$ where $\links(a)=v$ and $w$ names a $\1$-occurrence in $\Gamma$, the following are equivalent:
%
\begin{enumerate}
	\item
$\links\perm*\links'$ where $\links'(a)=w$;
	\item
$w$ is in the empire of $a$; and
	\item
in any switching graph for $(\Gamma,\links)$, the path from $v$ to $w$ does not pass through $a$.
\end{enumerate}
\end{lemma}

@@

Next, the empire of $a$ corresponds to the largest subproof $\Sigma$ in any $\Pi\toNet\links$ with as conclusion the introduction rule of $\named a\bot$.
%
By Definition~\ref{def:proofs to nets}, in the translation of $\Sigma$ to a net $\sigma$, $a$ may link anywhere in $\sigma$.
%
Then 2 implies 1.

Finally, if for some switching of $\links$ the path from $v$ to $w$ passes through $a$, then in $\links'$ there is no path between $a$ and $v$ (and two paths between $a$ and $w$) for that switching, so $\links'$ is not a net; then by contraposition 1 implies 3.
%
\end{IEEEproof}

\begin{lemma}
\label{lem:rewiring preserves empire}
Let $\links\perm*\links'$ by $\links(v)\neq\links'(v)$, and denote the empire of $v$ in $\links$ and $\links'$ by $\sigma$ and $\sigma'$ respectively; then $\sigma\perm*\sigma'$.
\end{lemma}

\begin{IEEEproof}
By Lemma~\ref{lem:rewiring within empire} the empires $\sigma$ and $\sigma'$ contain the same $\1$-subformulae, since $v$ may rewire to exactly the same $\1$-occurrences in $\links$ as in $\links'$.
%
That $\sigma$ and $\sigma'$ must share also any other subformula $A$ follows by the observation that $A\tn\1$ may replace $A$: by Proposition~\ref{prop:empire propagation} the new $\1$ is in a given empire if and only if $A$ is.
%
\end{IEEEproof}

\begin{lemma}
\label{lem:rewiring affects empires}
For proof nets $\links\perm*\links'$, let the empire of a given vertex $v$ in $\links$ be a net for the sequent $\Delta$, and in $\links'$, for $\Delta'$; then $\Delta\leq\Delta'$ or $\Delta\geq\Delta'$.
\end{lemma}

@@

\begin{lemma}
\label{lem:double exchange}
Let $\links$ be a proof net where $\links(a)=u$, $\links(b)=v$, $\links(c)=x$, and $\links(d)=y$, and for any switching the path from $u$ to $y$ passes through the links $\edge bv$ and $\edge xc$.
%
Then $\links\perm\links'$ where $\links'(a)=v$, $\links'(b)=u$, $\links'(c)=y$, and $\links'(d)=x$, and $\links'(z)=\links(z)$ otherwise.
\end{lemma}

@@

\begin{IEEEproof}
By induction on the size of a sequent $\Gamma$.
%
The base case is the sequent $\1,\1,\bot\tn\bot$, which has two inequivalent proof nets.
%
For the inductive step, let $A\tn\named a\bot$ be a largest $\tn$-formula in $\Gamma$, and pick an occurence $\named z\1$.
%
It will be shown that any net $\links$ is equivalent to one $\links'$ where $\links'(a)=z$; then by induction, for the sequent $\Gamma$ with $\named z\1$ removed and $A\tn\named a\bot$ replaced by $A$, the subnet for $\links'$ belongs to one of two equivalence classes.

To find $\links'$, if $\links(a)=z$ we are done.
%
Next, if the path from $\links(a)$ to $z$ does not pass through $a$, then by Lemma~\ref{lem:rewiring within empire} $\links'$ may be obtained from $\links$ by changing only $\links'(a)=z$.
%
Thirdly, let $\links(b)=z$ for a $\named b\bot$ in $A$.
%
Let $\named c\bot$ and $\named d\bot$ be occurrences in a separate formula $B$ such that $c$ links to the same $\1$-occurrence as some $\bot$ in $A$.
%
Then $\links'$ is obtained by first linking $c$ to $b$, and then applying Lemma~\ref{lem:double exchange} to exchange the targets of $a$ and $b$, as well as the targets of $c$ and $d$.

Finally, let the path from $a$ to $z$ pass $\named b\bot$ in $A$.
%
Then using Lemma~\ref{lem:rewiring within empire} re-attach $b$ to $z$, and repeat as above.
\end{IEEEproof}

@@

The standard linkings defined in the first and third case of the above definition are illustrated below.

>>>>

\author{
  \IEEEauthorblockN{Willem Heijltjes}
  \IEEEauthorblockA{
  		University of Bath
	%\\	Claverton Down
	%\\	Bath BA2~7AY
	\\	w.b.heijltjes@bath.ac.uk
	}
\and
  \IEEEauthorblockN{Robin Houston}
  \IEEEauthorblockA{
		Kiln Enterprises
	\\	robin@kiln.it
	}
}

@@

Later, a striking technical innovation of linear logic was the introduction of proof nets \cite{Girard-1987,Danos-Regnier-1989}, which define a canonical form for proofs in the unitless fragment of multiplicative linear logic -- two proofs are equivalent if and only if they have the same proof net -- so proof nets offer a simple decision procedure for proofs in this fragment. This naturally raises the question whether proof nets can be extended to work in the presence of units. The work in this direction begins with \cite{Trimble-1994,Blute-Cockett-Seely-Trimble-1996} via \color{red}\cite{Lamarche-Strassburger-2006}\color{black} and perhaps culminates in \cite{HughesFreeStar,HughesMLLProofNets}; but these proof nets are not canonical and must be identified up to a rewiring equivalence.

@@

In this paper we establish that the proof equivalence problem for multiplicative linear logic with units is \capsabbrev{pspace}-complete. It follows that there can be no satisfactory canonical notion of proof net: 
\color{red}
it is reasonable to ask that both the translation from proofs to nets and equality of proof nets are polynomial-time computable---but this would imply \capsabbrev{p}\,=\,\capsabbrev{pspace}.
\color{black}

@@

\begin{proposition}[\cite{Danos-Regnier-1989}, \color{red}\cite{Lamarche-Strassburger-2006}\color{black}]
\label{prop:correctness and sequentialisation}
%
%
If $\Pi\toNet\links$ and $\Pi$ has conclusion $\Gamma$, then $\links$ is a proof net for $\Gamma$.
%For a proof $\Pi$ with conclusion $\Gamma$, if $\Pi\toNet\links$ then $\links$ is a proof net for $\Gamma$.
%
For a net $\links$ for $\Gamma$, there is a proof $\Pi$ of $\Gamma$ such that $\Pi\toNet\links$ (\emph{sequentialisation}).
%
\end{proposition}

@@

We will denote by $A^n$ the sequent consisting of $n$ occurrences of a formula $A$.
%
Given a sequent $\Gamma=A_1,\dotsc,A_n$ we will write $\bigtn\Gamma$ for $A_1\tn\cdots\tn A_n$, and $\bigparr\Gamma$ for $A_1\parr\cdots\parr A_n$.
%
In diagrammatic notation, a big disc will represent an $n$-ary tensor over only $\bot$-formulae, $\bigtn(\bot^n)$, and a big circle will represent an $n$-ary par (or a sequent) over only $1$-formulae, $\bigparr(\1^n)$.
%
A linking between $\bigtn(\bot^n)$ and $\bigparr(\1^n)$ will be represented by a wide arrow, as illustrated below.
\[
\begin{tikzpicture}[x=-5mm,y=5mm,octo]
	\node[big_] (A) at (-1,2) {};
	\node[big1] (!) at (-1,0) {};
	\node at (.5,1) {$\Rightarrow$};
	\draw (2,2) node[bullet] (a) {} -- (3,2) node[bullet] (b) {} -- (4,2) (5,2) -- (6,2) node[bullet] (c) {}; 
	\node[circ] (1) at (2,0) {}; \node[circ] (2) at (3,0) {}; \node[circ] (3) at (6,0) {};
	\draw[dotted] (4,2)--(5,2) (4,0)--(5.1,0);
	\begin{scope}[thepink,->]
			\draw[big>] (A) to (!.center);
			\draw[bend right=10] (a) to (1.center);
			\draw[bend right=10] (b) to (2.center);
			\draw[bend right=10] (c) to (3.center);
	\end{scope}
\end{tikzpicture}
\]
Two formulae $\bigtn(\bot^{i+1})$ and $\bigtn(\bot^{j+1})$ may together connect to a formula $\bigparr(\1^{i+j+1})$, as illustrated below for $i=2$ and $j=3$.
%
Nodes will be labelled $i'$ for $i+1$, so that both the abbreviated formulae and the arithmetic of connecting them remain intuitive.
%
\[
\vc{
  \begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\draw (0,0) node[bullet] (a) {}--
		  (1,0) node[bullet] (b) {}--
		  (2,0) node[bullet] (c) {};
	\draw (4,0) node[bullet] (d) {}--
		  (5,0) node[bullet] (e) {}--
		  (6,0) node[bullet] (f) {}--
		  (7,0) node[bullet] (g) {};
	\foreach \i in {1,...,6} {\node[circ] (\i) at (\i,3) {};}
	\begin{scope}[thepink,->]
			\draw[bend right=10] (a) to (1);
			\draw[bend right=10] (b) to (2);
			\draw[bend right=10] (c) to (3);
			\draw[bend left=10]  (d) to (3);
			\draw[bend left=10]  (e) to (4);
			\draw[bend left=10]  (f) to (5);
			\draw[bend left=10]  (g) to (6);
	\end{scope}
  \end{tikzpicture}
}
\quad\Rightarrow\quad
\vc{
  \begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\node[big_] (A) at (-1,0) {$2'$};
	\node[big_] (B) at ( 1,0) {$3'$};
	\node[big1] (!) at ( 0,3) {$5'$};
	\begin{scope}[thepink,->,big>]
		\draw[bend right=15] (A) to (!);
		\draw[bend left=15]  (B) to (!);
	\end{scope}
  \end{tikzpicture}
}
\]
%
\color{red}
A path in a proof net or switching graph is indicated $\path ab$, and illustrated as below.
\color{black}
%
%A dashed edge will indicate that two vertices are connected, in a proof net or a switching graph, as illustrated below.
\[
\begin{tikzpicture}[x=5mm,y=5mm,octo]
	\node[bullet] (a) at (0,0) {}; \node[bullet] (b) at (2,0) {}; \draw[dashed] (a)--(b);
\end{tikzpicture}
\qquad
\begin{tikzpicture}[x=5mm,y=5mm,octo]
	\node[circ] (a) at (0,0) {}; \node[circ] (b) at (2,0) {}; \draw[dashed] (a)--(b);
	%\coordinate (a') at ($(a) + (.6,0)$); \coordinate (b') at ($(b) + (-.6,0)$);
	%\draw[thepink,dashed] (a')--(b'); \draw[thepink,->] (a')--(a.center); \draw[thepink,->] (b')--(b.center);
\end{tikzpicture}
\]

@@

The scopes of a par correspond to the possible subproofs of its introduction rule in a sequentialisation of the proof net that it occurs in.
%
\color{red}
In the graph of a proof net, the scope of a par can be made explicit as a \emph{box}, as illustrated below left.
%
Boxes may replace the switching condition as a correctness criterion: in the example, both the outside and the inside of the box form a tree.
%
To make this precise, we will consider the action of \emph{closing} a box, which will mean we consider it as a single vertex in the graph.

\[
\begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\draw (-2.25,0) node[bullet] (x) {} -- (-1.25,0) node[bullet] (y) {}
		  (.25,0) node[bullet] (a) {} -- (1.25,0) node[bullet] (b) {}
		  ( 2.75,0) node[bullet] (c) {} -- (3.75,0) node[bullet] (d) {};
	\node[circ] (0) at (5.75,0) {};
	\node[circ] (1) at (-1,3) {};
	\node[circ] (2) at (1,3) {}; \node[circ] (3) at (2,3) {}; \node[circ] (4) at (3,3) {};
	\draw[rounded corners] (0,2) rectangle (4,4);
	\draw (4,3) -- (5.5,3) node[bullet] (e) {};
	\begin{scope}[thepink,->]
			\draw[bend right=10] (x) to (1);
			\draw[bend right=20] (y) to (2);
			\draw[bend right=10] (a) to (2);
			\draw[bend right=10] (b) to (3);
			\draw[bend left=10]  (c) to (3);
			\draw[bend left=10]  (d) to (4);
			\draw[bend right=10] (e) to (0);
	\end{scope}
	\draw[dashed,rounded corners] (-.25,-.5) rectangle (4.25,4.25);
\end{tikzpicture}
\qquad\qquad
\begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\draw (-2.25,0) node[bullet] (x) {} -- (-1.25,0) node[bullet] (y) {};
	\node[circ] (0) at (3,0) {};
	\node[circ] (1) at (-1,3) {};
	\node[draw,rounded corners,minimum size=16pt] (box) at (1,3) {};
	\draw (box)--(3,3) node[bullet] (e) {};
	\begin{scope}[thepink,->]
			\draw[bend right=10] (x) to (1);
			\draw[bend right=20] (y) to (box);
			\draw[bend right=10] (e) to (0);
	\end{scope}
	\path (0,4.25)--(1,4.25);
\end{tikzpicture}
\]

\begin{definition}
A \emph{boxing} $s$ for a linking $\links$ for $\Gamma$ assigns a sub-sequent $s(v)\leq\Gamma$ to each par-vertex $\named v\parr$ such that 1) $v$ is a port of $s(v)$ and 2) boxes are either disjoint or strictly nested: if $s(v)\cap s(w)\neq\varnothing$ then $s(w)<s(v)$ or $s(v)<s(w)$.
\end{definition}

\noindent
In the graph for $\links$ and $\Gamma$, a box $s(v)$ may be \emph{closed} by replacing the subgraph over $s(v)$ by the single vertex $v$, and replacing every arc into $s(v)$ by one onto $v$.
%
For each box $s(v)$ we define the \emph{local graph} to be that formed by the subgraph over $s(v)$ where each immediately smaller box $s(w)<s(v)$ is closed.
%
The following is then a variation on the local retraction algorithm by Danos \cite{Danos-1990}.

\begin{proposition}
\label{prop:scoping correctness}
A linking $\links$ for $\Gamma$ is a proof net if and only if it has a boxing $s$ such that each local graph is a tree.
\end{proposition}

\begin{proof}
Given a boxing $s$, it follows by induction on the nesting of boxes that the graph over each $s(v)$ satisfies the switching condition.
%
In the other direction, given a sequentialisation of $(\Gamma,\links)$, a box $s(v)\leq\Gamma$ for each $\named v\parr$ is found by taking the conclusion $\Delta, A\named v\parr B$ of its introduction rule, below.
\[
	\infer[\MLLlabel p]{\Delta,A\named v\parr B}{\Delta,A,B}
\]
\vskip-12pt
\end{proof}

\color{black}

%
%In the graph of a proof net, the scope of a par $\named v\parr$ may be \emph{contracted} to a single vertex $v$ by removing all vertices except $v$,  and re-attaching the edges of any removed vertices to $v$.
%%
%In the example below, a scope of a (ternary) par is indicated (left) and contracted (right).
%\[
%\begin{tikzpicture}[x=5mm,y=-5mm,octo]
%	\draw (-2.25,0) node[bullet] (x) {} -- (-1.25,0) node[bullet] (y) {}
%		  (.25,0) node[bullet] (a) {} -- (1.25,0) node[bullet] (b) {}
%		  ( 2.75,0) node[bullet] (c) {} -- (3.75,0) node[bullet] (d) {};
%	\node[circ] (0) at (5.75,0) {};
%	\node[circ] (1) at (-1,3) {};
%	\node[circ] (2) at (1,3) {}; \node[circ] (3) at (2,3) {}; \node[circ] (4) at (3,3) {};
%	\draw[rounded corners] (0,2) rectangle (4,4);
%	\draw (4,3) -- (5.5,3) node[bullet] (e) {};
%	\begin{scope}[thepink,->]
%			\draw[bend right=10] (x) to (1);
%			\draw[bend right=20] (y) to (2);
%			\draw[bend right=10] (a) to (2);
%			\draw[bend right=10] (b) to (3);
%			\draw[bend left=10]  (c) to (3);
%			\draw[bend left=10]  (d) to (4);
%			\draw[bend right=10] (e) to (0);
%	\end{scope}
%	\draw[dashed,rounded corners] (-.25,-.5) rectangle (4.25,4.25);
%\end{tikzpicture}
%\qquad\qquad
%\begin{tikzpicture}[x=5mm,y=-5mm,octo]
%	\draw (-2.25,0) node[bullet] (x) {} -- (-1.25,0) node[bullet] (y) {};
%	\node[circ] (0) at (3,0) {};
%	\node[circ] (1) at (-1,3) {};
%	\node[draw,rounded corners,minimum size=16pt] (box) at (1,3) {};
%	\draw (box)--(3,3) node[bullet] (e) {};
%	\begin{scope}[thepink,->]
%			\draw[bend right=10] (x) to (1);
%			\draw[bend right=20] (y) to (box);
%			\draw[bend right=10] (e) to (0);
%	\end{scope}
%	\path (0,4.25)--(1,4.25);
%\end{tikzpicture}
%\]
%%
%The contraction of a scope corresponds to the use of a box,  may replace the switching condition as a correctness criterion:
%%
%The following is a variant of the local retraction algorithm by Danos \cite{Danos-1990}.

%
%\begin{proposition}
%\label{prop:scoping correctness}
%A linking $\links$ for a sequent $\Gamma$ is a proof net if and only if each $\named v\parr$ is a port of a sub-sequent $s(v)\leq\Gamma$ such that:
%
%\begin{enumerate}
%	\item
%sub-sequents either disjoint or strictly nested: if $s(v)\cap s(w)\neq\varnothing$ then $s(w)<s(v)$ or $s(v)<s(w)$; and
%	\item
%for each $\named v\parr$ the graph $\sigma(v)=(s(v),\links|_{s(v)})$ becomes a tree when all \color{red} (immediate) sub-sequents $s(w)<s(v)$ are contracted.
%\end{enumerate}
%%
%\end{proposition}
%
%
%\begin{IEEEproof}
%For the `if' direction, it follows by induction on the nesting of sub-sequents that each graph $\sigma(v)$ satisfies the switching condition.
%%
%For the `only if' direction, given a sequentialisation of $(\Gamma,\links)$, a sub-sequent $s(v)\leq\Gamma$ for each $\named v\parr$ is found by taking the conclusion $\Delta, A\named v\parr B$ of its introduction rule, below.
%\[
%	\infer[\MLLlabel p]{\Delta,A\named v\parr B}{\Delta,A,B}
%\]
%\vskip-12pt
%\end{IEEEproof}

In a proof net, the \emph{kingdom} and the \emph{empire} of a vertex $v$ are respectively the smallest and largest subnet that have $v$ as a port.
%
In working with the rewiring relation, the notion of empire can be particularly useful.
%
\color{red} 
We will denote the empire of $v$ in $\links$ by $\emp v$.
\color{black}

\begin{proposition}[{\cite[Proposition 2.b]{Bellin-vandeWiele-1995}}]
\label{prop:empire propagation}
The empire $\emp v$ is determined by propagation from $v$: 1) through links; 2) up towards subformulae; 3) into a tensor if one of its subformulae is in $\emp v$; 4) into a par if all its subformulae are in $\emp v$; 5) except never into a par or tensor if $v$ is a subformula.
\end{proposition}

@@

\begin{lemma}
\label{lem:rewiring within empire}
For a proof net $(\Gamma,\links)$ where $\links(a)=v$ and $w$ names a $\1$-occurrence in $\Gamma$, the following are equivalent:
%
\begin{enumerate}
	\item
$\links\perm*\links'$ where $\links'(a)=w$;
	\item
$w$ is in the empire $\emp a$; and
	\item
in any switching graph for $(\Gamma,\links)$, the path $\path vw$ does not pass through $a$.
\end{enumerate}
\end{lemma}

@@

Next, it is shown that 2 implies 1.
%
The empire $\emp a$ corresponds to the largest subproof $\Sigma$ in any $\Pi\toNet\links$ with as conclusion the introduction rule of $\named a\bot$.
%
By Definition~\ref{def:proofs to nets}, in the translation of $\Sigma$ to a net, $a$ may link to any $1$-occurrence, including to $w$.

Finally, it is shown that 1 implies 3, by contraposition.
%
If for some switching of $\links$ the path $\path vw$ passes through $a$, then in $\links'$ there is no path $\path av$ (and two paths $\path aw$) for that switching, so $\links'$ is not a net.
%
\end{IEEEproof}

\begin{lemma}
\label{lem:rewiring preserves empire}
If $\links\perm*\links'$ by $\links(v)\neq\links'(v)$ then $\emp v\perm*\emp*v$.
\end{lemma}

\begin{IEEEproof}
By Lemma~\ref{lem:rewiring within empire} the empires $\emp v$ and $\emp*v$ contain the same $\1$-subformulae, since $v$ may rewire to exactly the same $\1$-occurrences in $\links$ as in $\links'$.
%
That $\emp v$ and $\emp*v$ must share also any other subformula $A$ follows by the observation that $A\tn\1$ may replace $A$: by Proposition~\ref{prop:empire propagation} the new $\1$ is in a given empire if and only if $A$ is.
%
\end{IEEEproof}

\begin{lemma}
\label{lem:rewiring affects empires}
If $\links\perm*\links'$ where $\emp v$ is a net for the sequent $\Delta$, and $\emp*v$ a net for $\Delta'$, then $\Delta\leq\Delta'$ or $\Delta\geq\Delta'$.
\end{lemma}

@@

\begin{lemma}
\label{lem:double exchange}
Let $\links$ be a proof net with for every switching a path
$\links(a){\xedge}a{\xpath}b{\xedge}\links(b){\xpath}\links(c){\xedge}c{\xpath}d{\xedge}\links(d)$.
%
Then $\links\perm\links'$ where $\links'(a)=\links(b)$ and $\links'(b)=\links(a)$, $\links'(c)=\links(d)$, and $\links'(d)=\links(c)$, and $\links'(z)=\links(z)$ otherwise.
%
%Let $\links$ be a proof net where $\links(a)=u$, $\links(b)=v$, $\links(c)=x$, and $\links(d)=y$, and for any switching the path $\path uy$ passes through $\edge bv$ and $\edge xc$.
%
%Then $\links\perm\links'$ where $\links'(a)=v$, $\links'(b)=u$, $\links'(c)=y$, and $\links'(d)=x$, and $\links'(z)=\links(z)$ otherwise.
\end{lemma}

@@

\begin{IEEEproof}
By induction on the size of a sequent $\Gamma$.
%
The base case is the sequent $\1,\1,\bot\tn\bot$, which has two inequivalent proof nets.
%
For the inductive step,
%
\color{red}
let $\Gamma=\Gamma',A\tn\named a\bot,\named z\1$ where $A\tn\named a\bot$ is a largest $\tn$-formula in $\Gamma$.
%
It will be shown that any net $\links$ is equivalent to one $\links'$ where $\links'(a)=z$; then by induction, the subnet $\links'$ restricted to $\Gamma',A$ belongs to one of two equivalence classes.
%
To find $\links'$, there are two cases.

1) The path $\path az$ is via $\links(a)$.
%
If $\links(a)=z$, we are done. Otherwise, by Lemma~\ref{lem:rewiring within empire} $\links'$ may be obtained from $\links$ by changing only $\links'(a)=z$.

2) The path $\path az$ is via some $\named b\bot$ in $A$.
%
Firstly, if $\links(b)\neq z$, use Lemma~\ref{lem:rewiring within empire} to re-attach $b$ to $z$.
%
Next, let $\named c\bot$ and $\named d\bot$ be occurrences in a separate formula $B$ such that $c$ links to the same $\1$-occurrence as some $\bot$ in $A$.
%
Then $\links'$ is obtained by linking $c$ to $b$, and applying Lemma~\ref{lem:double exchange} to exchange the targets of $a$ and $b$, as well as those of $c$ and $d$.
%
\end{IEEEproof}
\color{black}

%
%To find $\links'$, if $\links(a)=z$ we are done.
%%
%Next, if the path from $\links(a)$ to $z$ does not pass through $a$, then by Lemma~\ref{lem:rewiring within empire} $\links'$ may be obtained from $\links$ by changing only $\links'(a)=z$.
%%
%Thirdly, let $\links(b)=z$ for a $\named b\bot$ in $A$.
%%
%Let $\named c\bot$ and $\named d\bot$ be occurrences in a separate formula $B$ such that $c$ links to the same $\1$-occurrence as some $\bot$ in $A$.
%%
%Then $\links'$ is obtained by first linking $c$ to $b$, and then applying Lemma~\ref{lem:double exchange} to exchange the targets of $a$ and $b$, as well as the targets of $c$ and $d$.
%
%
%Finally, let the path from $a$ to $z$ pass $\named b\bot$ in $A$.
%%
%Then using Lemma~\ref{lem:rewiring within empire} re-attach $b$ to $z$, and repeat as above.

@@

The standard linkings defined in the \color{red} second \color{black} and third case of the above definition are illustrated below.
